Leadership across cultures and contexts: an exploratory analysis

### 2020-05-02 leadership_across_cultures_contexts.Rmd

The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Supplimentary Figure \@ref(fig:subsisttree)). Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures. The number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`. The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`. Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one variable in the study.

@@

Subsistence strategy was confounded with continental region. African text records were predominantly from horticultural societies, for instance, whereas North American text records were predominantly from hunter-gatherer societies (see Supplimentary Figure \@ref(fig:regionsubsistence)).

@@

We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` variables in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)). Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Supplimentary Figure \@ref(fig:doc-years)). There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors). Of the `r nrow(documents)` source documents, `r sum(documents$female_coauthor)` (`r signif(100*sum(documents$female_coauthor)/nrow(documents), 2)`%) had a female author or co-author. Total pages of ethnography on each culture in the PSF ranged from `r min(leader_cult[['ehraf_pages']])` to `r as.integer(max(leader_cult[['ehraf_pages']]))`, with a median of `r median(leader_cult[['ehraf_pages']])`.

The number of text records on leadership retrieved for a particular culture was weakly correlated with the total number of pages of ethnography in the eHRAF (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`). A loess regression suggested that the number of text records initially increased with the total number of pages of ethnography, plateauing at about 35 records around 4500 pages (see Supplimentary Figure \@ref(fig:eHRAFpages)).

We fit `r length(all_study_vars)` logistic mixed effects models with each binary study variable as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures. We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995]. After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but the publication date coefficient was significant in nine models. More recent publications were more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*). See Supplimentary Figure \@ref(fig:pubdate) and supplementary information for details.

@@

*High status* was the most frequently identified leader quality in the sample, with `r statusPercTxt`% of text records and `r statusPercCult`% of cultures containing evidence that leaders were of high social status. To explore factors that distinguished these records from those with no evidence for high status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occurred in each text record (see supplimentary information for details). We then fit an elasticnet logistic regression model (with lasso penalty) of *High status* (0/1) as a function of the frequencies of all `r ncol(leader_dtm)-1` words. Words that were strong positive predictors epitomized the semantic content of the text records which provided evidence of high status as a quality of leaders. These included *`r rev(highstatus_plot$data$Word)[1]`* and *`r rev(highstatus_plot$data$Word)[2]`*, which were the strongest predictive terms; *`r rev(highstatus_plot$data$Word)[3]`* and *`r rev(highstatus_plot$data$Word)[4]`* were also relatively strong predictors and *`r rev(highstatus_plot$data$Word)[5]`* and *`r rev(highstatus_plot$data$Word)[6]`*, which were also positive predictors. Notable terms which offered predictive value but with near-zero coefficients included, *influence*, *authority*, *headman*, and *chief*. See Supplimentary Figure \@ref(fig:highstatusplot).

@@

Here we report the cluster analysis of variables that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from the other methods, see the supplimentary information). Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables. Each cluster analysis includes two estimates of significance for how strongly each cluster within the dendrogram is supported by the data. We rely on the AU (Approximately Unbiased) *p* values, which are computed by multiscale bootstrap resampling and represented as percentages (e.g., clusters with AU values $>95$ are strongly supported, and the top-level clusters are automatically highlighted by rectangles).

>>>>

The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Supplementary Figure \@ref(fig:subsisttree)). Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures. The number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`. The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`. Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one variable in the study.

@@

Subsistence strategy was confounded with continental region. African text records were predominantly from horticultural societies, for instance, whereas North American text records were predominantly from hunter-gatherer societies (see Supplementary Figure \@ref(fig:regionsubsistence)).

@@

We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` variables in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)). Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Supplementary Figure \@ref(fig:doc-years)). There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors). Of the `r nrow(documents)` source documents, `r sum(documents$female_coauthor)` (`r signif(100*sum(documents$female_coauthor)/nrow(documents), 2)`%) had a female author or co-author. Total pages of ethnography on each culture in the PSF ranged from `r min(leader_cult[['ehraf_pages']])` to `r as.integer(max(leader_cult[['ehraf_pages']]))`, with a median of `r median(leader_cult[['ehraf_pages']])`.

The number of text records on leadership retrieved for a particular culture was weakly correlated with the total number of pages of ethnography in the eHRAF (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`). A loess regression suggested that the number of text records initially increased with the total number of pages of ethnography, plateauing at about 35 records around 4500 pages (see Supplementary Figure \@ref(fig:eHRAFpages)).

We fit `r length(all_study_vars)` logistic mixed effects models with each binary study variable as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures. We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995]. After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but the publication date coefficient was significant in nine models. More recent publications were more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*). See Supplementary Figure \@ref(fig:pubdate) and supplementary information for details.

@@

*High status* was the most frequently identified leader quality in the sample, with `r statusPercTxt`% of text records and `r statusPercCult`% of cultures containing evidence that leaders were of high social status. To explore factors that distinguished these records from those with no evidence for high status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occurred in each text record (see supplementary information for details). We then fit an elasticnet logistic regression model (with lasso penalty) of *High status* (0/1) as a function of the frequencies of all `r ncol(leader_dtm)-1` words. Words that were strong positive predictors epitomized the semantic content of the text records which provided evidence of high status as a quality of leaders. These included *`r rev(highstatus_plot$data$Word)[1]`* and *`r rev(highstatus_plot$data$Word)[2]`*, which were the strongest predictive terms; *`r rev(highstatus_plot$data$Word)[3]`* and *`r rev(highstatus_plot$data$Word)[4]`* were also relatively strong predictors and *`r rev(highstatus_plot$data$Word)[5]`* and *`r rev(highstatus_plot$data$Word)[6]`*, which were also positive predictors. Notable terms which offered predictive value but with near-zero coefficients included, *influence*, *authority*, *headman*, and *chief*. See Supplementary Figure \@ref(fig:highstatusplot).

@@

Here we report the cluster analysis of variables that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from the other methods, see the supplementary information). Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables. Each cluster analysis includes two estimates of significance for how strongly each cluster within the dendrogram is supported by the data. We rely on the AU (Approximately Unbiased) *p* values, which are computed by multiscale bootstrap resampling and represented as percentages (e.g., clusters with AU values $>95$ are strongly supported, and the top-level clusters are automatically highlighted by rectangles).
